---
title: 'CWRU DSCI351-351m-451: Lab Exercise LE2'
author: 'Prof.:Roger French, TA: Raymond Wieser, Sameera Nalin Venkat, Mingxuan Li'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
  word_document:
    toc: yes
    toc_depth: '6'
  pdf_document:
    latex_engine: xelatex
    toc: yes
    number_sections: yes
    toc_depth: 6
    highlight: tango
subtitle: Timeseries Analysis, ggplot2, Text Mining
urlcolor: blue
---

<!--
# Script Name: 2008-351-351m-451-LE2-NAME.Rmd
# Purpose: This is an Rmd LE file to do your LE in
# Authors: Roger H. French
# License: All rights reserved.
##########
# Latest Changelog Entires:
# v0.00.01 - 1501cwru-dsci-NAMEIT.RMD - Roger French started this blank Rmd
##########

# Rmd code goes below the comment marker!
-->

\setcounter{section}{2}
\setcounter{subsection}{0}

### LE2, 7 points, 3 questions. 

Summary of points (use Cntrl + Shift + O for seeing sub-questions easily):-

LE2-1: 2,5 points

  - LE2-1a: 1 point
  - LE2-1b: 1.5 points
  
LE2-2: 1.5 points

  - LE2-2a, LE2-2b: 0.25 points each
  - LE2-2c, LE2-2d: 0.5 points each
  
LE2-3: 3 points

  - LE2-3a, LE2-3b: 0.5 point each
  - LE2-3c: 1 point
  - LE2-3d: 0.5 point
  - LE2-3e: 1 point


#### Lab Exercise (LE) 2

--------------------------------------------

## LE2-1. Time series analysis: (2.5 points)

Time series are a common type of data, 

  - consisting of measurements that are continuous over a time range. 

In this project we will be using classical decomposition 

  - to perform analysis on a time series.

First as an introduction to decomposition we will have a quick example.

### LE2-1a (1 point)

  - What is the decomposition of a time-series?
  
Answer - It is a way by which we can analyse a data over time.
Time series decomposition involves thinking of a series as a combination of level, trend,          seasonality, and noise components.
Decomposition provides a useful abstract model for thinking about time series generally and for   better understanding problems during time series analysis and forecasting.
  
  - The AirPassengers data set of airline passengers every month for 12 years


```{r, include = FALSE}
# Load in required libraries, readr for csv files
library(readr)
library(tidyverse)
library(datasets)
```

```{r}


## playground
# data(AirPassengers)
# is.ts(AirPassengers)
# class(AirPassengers)
# start(AirPassengers)
# end(AirPassengers)
# frequency(AirPassengers)
# boxplot(AirPassengers ~ cycle(AirPassengers))
# plot(diff(log(AirPassengers)))
# AP <- AirPassengers
# View(AP)

# data.frame(as.integer(rownames(AirPassengers)))

# playground ends
```




```{r}

data(AirPassengers) #Loads specified data sets, or list the available data sets.
AirPassengers
AirPassengers <- as.data.frame(AirPassengers)   #converting a ts object to a dataframe ;this means # # loosing out on all the timeinformation
is.data.frame(AirPassengers)
# head(AirPassengers)

```

  - Plot the total time series of air passengers 
  - What do you notice about the plot?
  
  

```{r}
  plot(AirPassengers)
  # length(AirPassengers)
  # frequency(AirPassengers)
  
  # 
  # abline(reg = lm(AP ~ time(AP)))
  # plot(aggregate(AP,FUN =mean))
  # 
  # summary(AirPassengers)
```
ANSWER (what do you notice about the plot?) -> There is a periodicity in every year, and the amplitude keeps on increasing every year.

  - Use the ts() function in base R 
    - to define AirPassengers as a time series with a yearly trend 
  - If the data is taken monthly, 
    - what will the frequency (points per season) of a yearly season be? 
    
    

```{r}
# Create a time series object with a yearly trend
?ts

ap_y <- ts(data = AirPassengers, frequency = 12)
ap_y


```
ANSWER (what will the frequency (points per season) of a yearly season be? ) ->
if there are 4 seasons in a year of 12 months, then each season will have 3 points



  - Use the decompose() function 
    -to decompose the time series and remove the seasonality
  - The type for this time series is multiplicative
  - Plot the decomposed time series, what do you notice about the trend?
  
  

```{r}
# playground
?decompose

# View after decomposition into base components

ap_y_d <-decompose(ap_y,type = "multiplicative")
plot(ap_y_d$x)
plot(ap_y_d$trend)
plot(ap_y_d$seasonal)
# ap_y_d2 <- (ap_y / ap_y_d$seasonal)   #is this correct ?
# plot(ap_y_d2)



```

```{r}

# Removing seasonality


ap2 <-decompose(ap_y, type = "multiplicative")
ap3 <- ap_y /ap2$seasonal # removing seasonal

plot(ap3)



```
ANSWER (what do you notice about the trend?) ->
  The trend of passengers keeps on increasing as the years pass.

  - Isolate the trend and plot the trend on top of the raw data with the seasonality included
  - How well does the trend represent the data?

```{r}
# Plotting simulated valued that is trend and seasonal vs actual

simulated <- (ap2$trend * ap2$seasonal)
plot( simulated, type="l", col="red" )
lines(ap_y, col="green")


```
ANSWER (how well does the trend represent the data?) ->

The trend very closely follows the actual data

### LE2-1b (1.5 points)

Now lets try this with a real world time series. 
We'll be using one month of power and weather data from a solar power plant.

The data set variables are as follows:

  - time: The timestamp at which the data was taken
  - ghir: Global Horizontal Irradiance from a sensor at the site, 
    - the power from the sunlight over an area normal to the surface of the earth $(Watts/m^2)$
  - iacp: The AC power from the power plant $(kW)$
  - temp: The air temperature $(Celsius)$ 
  - ghi_solargis: The Global Horizontal Irradiance, not from a sensor, 
    - but predicted using weather modeling $(Watts/m^2)$
  - clear: A logical indicating whether the sky was "clear" during measurement, 
    - determined by comparing the ghi and ghi_solargis data
  - ratio: the ratio of the Global Horizontal Irradiance 
    - and the Plane of Array Irradiance (the irradiance normal to the surface of a tilted module)

The power from solar panels is dependent on the irradiance hitting it, 

  - so a power reading is often meaningless without a corresponding irradiance measurement.

It is useful to have multiple sources of irradiance measurements. 

Sensors on the ground are useful because 

  - they strongly represent the irradiance that is hitting the module; 
  - however, sensors can begin to drift if not cleaned or calibrated properly. 
  - An unstable sensor can render an entire data set useless.

To combat this, we also have irradiance data from SolarGIS, 

  - a company that uses satellite images to model and predict 
  - the irradiance at the surface of the earth. 

  - Plot the irradiance and power for the first week of data, 
    - how does the irradiance look compared to 
      - what you would expect from the trend of sunlight? 
    - How well does the power and irradiance match up?

```{r}

# input1 <-read.csv("data/2108-351-351m-451-LE1-solectria-data-final.csv", header=TRUE)
# 
# is.ts(input1) # check if it is a timeseries
# summary(input1)

# plot(input1$ghi_solargis,type = "l")
# plot(input1$iacp, type ="l")
# plot(input1$ghir, type = "l")




```



```{r}
######################################################
# Read in the data and get a look at the structure
input1 <-read.csv("data/2108-351-351m-451-LE1-solectria-data-final.csv", header=TRUE)
head(input1)

######################################################
# Find total length of dataset
dim(input1)[0]

######################################################
# Find the time interval
difftime(input1$time[2], input1$time[1],
         units = "mins")

######################################################
# Plot the irradiance and power for the first week
class(input1$time)
# as.character(input1$time)
x <- input1$time
# y <- format(input1$time,  "%X" )
# z <- strptime(x ,format = "", tz = "")
z <- as.Date(x, "%Y-%m-%d")
input1$date <-z
# head(input1)

# inputf1 -> input1[]
dt_fst_wk <- subset(input1, date>="2012-06-01" & date<="2012-06-07")








##############


# Entire dataset is in 15-minute increment

# Power plot over time
timedt <- dt_fst_wk$time
pwr <- dt_fst_wk$iacp

plot(timedt,  pwr , col="blue", type ="l",main = "power vs time" ,pch="o", lty=1 )

# Both sources of irradiance
sen_ghi <- dt_fst_wk$ghir
gis_ghi <- dt_fst_wk$ghi_solargis

plot(timedt,  sen_ghi , col="green", type ="l", main="Sources of Irradiance",
ylab=" sensor in black and solargis in red" )
lines(timedt, gis_ghi , col="red", type ="l")








  
```

```{r}

# #### Playground
# ## locale-specific version of date()
# format(Sys.time(), " %a %b %d %X %Y %Z")
# # "Mon Sep 13 11:02:51 AM 2021 EDT"
# 
# ## time to sub-second accuracy (if supported by the OS)
# format(Sys.time(), "%H:%M:%OS3")
# 
# (x <- strptime(c("2006-01-08 10:07:52", "2006-08-07 19:33:02"),
#                "%Y-%m-%d %H:%M:%S", tz = "EST5EDT"))
# attr(x, "tzone")
# 
# ## read in date info in format 'ddmmmyyyy'
# ## This will give NA(s) in some non-English locales; setting the C locale
# ## as in the commented lines will overcome this on most systems.
# ## lct <- Sys.getlocale("LC_TIME"); Sys.setlocale("LC_TIME", "C")
# x <- c("1jan1960", "2jan1960", "31mar1960", "30jul1960")
# z <- strptime(x, "%d%b%Y")
# ## Sys.setlocale("LC_TIME", lct)
# class(z)
# x <-"2012-06-01 00:00:00"
# as.character(x)
# y <- strptime(x, format = "", tz = "")
# class(y)

```
ANSWER (how does the irradiance look compared to what you would expect from the trend of sunlight?) -> The irradiance captured by the sensor closely follows the solar gis values. 

ANSWER (how well does the power and irradiance match up?) ->
        The power and irradiance have the same period and they match up well.

```{r}
# All 3 lines on the same plot with a secondary axis

ggplot() +
  geom_line(aes(x= timedt, y = sen_ghi), group =1, color ="green")+
  geom_line(aes(x= timedt, y = gis_ghi), group =1, color ="blue")+
  geom_line(aes(x= timedt, y = pwr), group =1, color ="red")+
  scale_y_continuous(sec.axis = sec_axis(trans =~./10, name ="pwr"))
  


# # par(mar = c(5, 4, 4, 4) + 0.3)              # Additional space for second y-axis
# plot(timedt, pwr, pch = 16, col = "blue")              # Create first plot
# par(new = TRUE)                             # Add new plot
# plot(timedt, sen_ghi, pch = 17, col = "green",              # Create second plot without axes
#      axes = FALSE, xlab = "", ylab = "")
# lines(timedt, gis_ghi , col="red", type ="l")
# axis(side = 4, at = pretty(range(sen_ghi)))      # Add second axis
# mtext("sen_ghi", side = 4, line = 3)


```

- Use the ts() functions and the stlplus() function from the stlplus package 

  - to decompose the sensor and SolarGIS irradiance and the power 
    - for the whole month. 
  -  Plot each of the decompositions, what do you notice?

```{r}
# think carefully about the frequency you'll need to define for this data
# what is the seasonal component to this data and how many data points make up a season?
# use s.window = "periodic" for the stlplus function
library(stlplus)
?stlplus()


#convert data to timeseries objects
sen_ghi <- ts(data = input1$ghir,  frequency =4*24) ## change this
gis_ghi <- ts(data = input1$ghi_solargis , frequency = 4*24)
pwr <- ts(data = input1$iacp, frequency = 4*24)

?
sen_ghi

#decompose using stlplus
sg_d <- stlplus(sen_ghi,s.window = "periodic")
gg_d <- stlplus(gis_ghi,s.window = "periodic")
pwr_d <- stlplus(pwr,s.window = "periodic")


#plots1 - 4
plot(sg_d$data$raw, type="l")
plot(sg_d$data$seasonal, type="l")
plot(sg_d$data$trend, type="l")
plot(sg_d$data$remainder, type="l")

#plots5-8
plot(gg_d$data$raw, type="l")
plot(gg_d$data$seasonal, type="l")
plot(gg_d$data$trend, type="l")
plot(gg_d$data$remainder, type="l")


#plots5-8
plot(pwr_d$data$raw, type="l")
plot(pwr_d$data$seasonal, type="l")
plot(pwr_d$data$trend, type="l")
plot(pwr_d$data$remainder, type="l")





```
ANSWER (what do you notice about the decompositions?) ->

  - Isolate the trends for the 3 time series you just decomposed 
    - and build a linear model for each one.
  - Compare the models to each other, how are they different?

```{r}

y <- sg_d$data$trend
x <- sg_d$time
plot(sg_d$data$trend, type="l")

model1 <- lm(y ~ x)
plot(x,y)
abline(model1, lwd =3, col ="red")

summary(model1)

y <- gg_d$data$trend
x <- gg_d$time
plot(gg_d$data$trend, type="l")

model2 <- lm(y ~ x)
plot(x,y)
abline(model2, lwd =3, col ="red")

summary(model2)

y <- pwr_d$data$trend
x <- pwr_d$time
plot(pwr_d$data$trend, type="l")

model3 <- lm(y ~ x)
plot(x,y)
abline(model3, lwd =3, col ="red")

summary(model3)

```


ANSWER (compare the models to each other and how are they different?) ->
      both sensor sunlight data and Gis show an increase in the years with almost the same slope, even though power has increased over the years, it has risen slowly.

  - Solar panel degradation leads to less power output over time 
    - at the same irradiance conditions.
  - Based on the linear models you found for the trends of power and irradiance, 
    - is this system degrading over time?
  - How do the sensor GHI and the SolarGIS GHI compare to power?
  
ANSWER (is this system degrading over time based on linear models?) ->
        Yes, since power has risen slower than the growth rate of irradiance, it is likely that the system has degraded over time.

ANSWER (how do the sensor GHI and the SolarGIS GHI compare to power?) ->
        The sensor GHI and Solar GIS GHI, have grown the same way, it is likely that the sensors             are working well.

--------------------------------------------

## LE2-2. ggplot2: (1.5 points)

ggplot2 is a package for making plots from data. 

It provides tools for making complex and detailed graphs. 

ggplot2 builds graphs in layers, 

  - where first the graph must be specified, 
  - then layers are added to the plot using the '+' operator.

In this example nothing appears in the plot

```{r}
library(ggplot2)
data("mpg")
head(mpg)

ggplot(data = mpg, aes(x = cyl, y = hwy))

```

This is because we did not define the next layer, 

  - all we did was define some kind of graph between cylinders and highway mpg

Since we have already defined the data at the beginning, 

  - we don't need to specify it in the layer

```{r}

ggplot(data = mpg, aes(x = cyl, y = hwy)) +
  geom_point()

# or a different layer

# here we have to define cyl as the group for each box
ggplot(data = mpg, aes(x = cyl, y = hwy)) +
  geom_boxplot(aes(group = cyl))

```

We can add additional information about showing data in our plot 

  - by adding parameters into the aesthetics (aes()) function

```{r}

ggplot(data = mpg, aes(x = cty, y = hwy)) +
  geom_point(aes(size = displ))

```

We can also add on additional layers if we want to, 

  - keep in mind ordering is important. 
  
The data for each layer can be defined per layer 

  - this is important if you're trying to add multiple data sets to a plot

```{r}

ggplot() +
  geom_point(data = mpg, aes(x = cty, y = hwy)) +
  geom_smooth(data = mpg, aes(x = cty, y = hwy), method = "lm")

```

Now it's your turn to make some plots

  - All plotting must be done using ggplot2, 
  - Any data manipulation must be done with dplyr pipelines 
    - running into the ggplot funciton

### LE2-2a (0.25 points) 
2a: Use the mtcars data set, 
  
  - plot mpg vs displacement and color by cylinders
 
```{r}

data("mtcars")
head(mtcars)
ggplot(data = mtcars, aes(x = mpg, y = disp)) + geom_point(aes(color = as.factor(cyl)))

?mtcars



```

### LE2-2b (0.25 points)
2b: Create a boxplot of the horsepower readings for each cylinder count, 

  - show the data points on top of the plot
 
```{r}
ggplot(data = mtcars, aes(x = cyl, y = hp)) +
  geom_boxplot(aes(group = cyl))

```

### LE2-2c (0.5 points)
2c: Plot a histogram of the number cars in each carburetor count group

```{r}
ggplot(mtcars, aes(carb)) +
  geom_histogram(binwidth =1)
```

### LE2-2d (0.5 points)
2d: Explain why these two plots look different, 

  - why does the color and key change between them?
 
```{r}

ggplot(data = mpg, aes(x = cty, y = hwy)) +
  geom_point(aes(color = cyl, size = displ))

ggplot(data = mpg, aes(x = cty, y = hwy)) +
  geom_point(aes(color = factor(cyl), size = displ))

```

ANSWER (explain why these two plots look different) ->
plot 1 : Shows gradual change in the cty - indicating it as a continuous value
plot 2 : treats cty as a discreet values and gives each value of cty a different color to emphasise demarcation

ANSWER (why does the color and key change between them?) ->
The first one is more for a continuos cty and the second one is for a discreet cty

--------------------------------------------

## LE2-3. Text Mining of Song Lyrics: (3 points)

- Complete the given problems
- dplyr, ggplot, and pipes and pipelines are highly recommended
- We will be using the Tidytext package to aid with our text mining
- The dataset for this assignment is a collection of the information and lyrics from every top 100 billboard song since 1965

### LE2-3a (0.5 points)

- Load in the data set
- Print the lyrics of the #4 song from 1988
- Show the top 20 artists with the most hits, which artist has the most total top 100 hits?

```{r}
################################################ load in data
musi_inp1 = read.csv("data/billboard_lyrics_1964-2015.csv")
# View(musi_inp1)
###################################################\
library(dplyr)
sr <-musi_inp1[musi_inp1$Rank == 4 & musi_inp1$Year == 1988,]
# filter(musi_inp1, Rank == 4 , Year == 1988)
sr$Lyrics

###############################################

music <- tbl_df(musi_inp1)

top_artist <-music %>%
  group_by(Artist) %>%
  tally()%>%
  arrange(desc(n))

top_artist$Artist[1:20]
top_artist$Artist[1]




```

ANSWER -> The artist that has the most total top 100 hits is ___madonna________

### LE2-3b (0.5 points)

- Build a histogram of the amount of times artists appear on the top hits billboard, what does the trend look like, what does it suggest about 1 hit wonders?

ANSWER -> There are many artists, more than 1500, who appear very few times in the top 100 hits.
There is only 1 artist who has appeared upto 35 times.

```{r}


hitsvsartist <- top_artist %>%
  group_by(n)%>%
  tally(name = "n2")
hitsvsartist
  
  
ggplot(top_artist, aes(n)) +
  geom_histogram(binwidth =1)


```

### LE2-3c (1 point)

- Lets do a lyrical comparison between 2 top artists, Elton John and Eminem
- First, filter the main data set to only the 2 given artists (you can make them seperate data frames or keep them together, your call)
- Use the unnest_tokens() function from tidytext to split the lyrics up so that each word has its own row
- Show the top 10 most commonly used words for each artist

```{r}

# convert factor to character
music$Lyrics <- as.character(music$Lyrics)


mus_ej <- music %>%
  filter(Artist =="elton john" )
head(mus_ej)

mus_emnm <- music %>%
  filter(Artist =="eminem")
head(mus_emnm)



# break lyrics to words
ej_words <- mus_ej %>%
  unnest_tokens(output = word, input = Lyrics)

head(ej_words)


emnm_words <- mus_emnm %>%
  unnest_tokens(output = word, input = Lyrics)

head(emnm_words)

# top 10 words for each artist
ej_top_10 <- ej_words %>%
  group_by(word)%>%
  tally(name = "word_count")%>%
  arrange(desc(word_count))

print("Elton john uses thesa a lot")
ej_top_10$word[1:10]

em_top_10 <- emnm_words %>%
  group_by(word)%>%
  tally(name = "word_count")%>%
  arrange(desc(word_count))

print("eminem uses thesa a lot")
em_top_10$word[1:10]
  

```

- It should be clear that most of these are not significantly meaningful words and should be removed, we can remove them using the stop_words dataframe provided with the tidytext package (hint: look at dplyr anti_join())

- Remove them and show the top ten remaining words for each artist

```{r}

data("stop_words")

usable_ej_10 <- ej_top_10 %>%
  anti_join(stop_words, by = "word")

print("elton john uses these words most")
usable_ej_10$word[1:10]
  

print("eminem uses these words most")
usable_em_10 <- em_top_10%>%
  anti_join(stop_words, by = "word")

usable_em_10$word[1:10]



```

### LE2-3d (0.5 points)

- Build a word cloud of the lyrics for each artist, find an R package that will help you with this
- Compare and contrast the word clouds
- Did the stop_words dataframe work well to remove non-meaningful words?

```{r}

library(wordcloud)

usable_ej <- ej_words %>%
  anti_join(stop_words, by = "word")
usable_em <- emnm_words %>%
  anti_join(stop_words, by = "word")

wordcloud(words = usable_ej$word,  min.freq = 1,           max.words=200, random.order=FALSE, rot.per=0.35,            colors=brewer.pal(8, "Dark2"))
wordcloud(words = ej_words$word,  min.freq = 1,           max.words=200, random.order=FALSE, rot.per=0.35,            colors=brewer.pal(8, "Dark2"))

wordcloud(words = usable_em$word,  min.freq = 1,           max.words=200, random.order=FALSE, rot.per=0.35,            colors=brewer.pal(8, "Dark2"))
wordcloud(words = emnm_words$word,  min.freq = 1,           max.words=200, random.order=FALSE, rot.per=0.35,            colors=brewer.pal(8, "Dark2"))


# wordcloud(words = usable_em_10$word, max.words = 10 )



```

ANSWER (compare and contrast the word clouds) ->
Both Eminem and Elton john use "don't" a lot. 

ANSWER (did stop_words work well to remove non-meaningful words?) ->
The stop words well to remove non- meaningful words.
Even though word cloud tries to remove non- meanigful words, in the case of eminem, if we use his raw words without stop words filter a lot of  non-meaningful words like "get, he's, yeah " appear

### LE2-3e (1 point)

- The tidytext package gives us the ability to run sentiment analysis as well
- There are multiple sentiment methods available but let's just use positive and negative sentiment

```{r}
# library(tidytxt)
bing <- get_sentiments("bing")
head(bing)



```

- What is the bing dataframe?

ANSWER (definition of bing dataframe) -> 
  The bing dataframe : is a data set of many words  witha a sentiment attached to each word.

- Pull out the "positive" and "negative" words for each artist (hint: dplyr inner_join())
- For each song per artist, determine the next positive or negative sentiment, there are several ways to do this - the most straightforward being assign a 1 to a positive lyric and -1 to a negtive lyric then sum them together for each song
- Make a bar plot of the net sentiment of each song for each artist, make these plots high quality as well
  - Properly name the axes and title each plot
  - Color by whether the song is overall positive or negative
  - rotate the x-axis names so long song titles are legible
  - Arrange the bars in ascending or descending order

```{r}

## words used by each artist, and assigning it with a sentiment using bing
usable_ej_sent <-  merge(x = usable_ej, y = bing, by = "word")
usable_ej_sent <- usable_ej_sent %>%
  mutate(sen_value = if_else( usable_ej_sent$sentiment == "negative", -1 , 1, 0))

usable_ej_sent <-usable_ej_sent %>% 
  group_by(Song)%>%
  summarise(net_Sen = sum(sen_value))

usable_ej_sent <- usable_ej_sent %>%
  arrange(desc(net_Sen))

usable_em_sent <-  merge(x = usable_em, y = bing, by = "word")
usable_em_sent <- usable_em_sent %>%
  mutate(sen_value = if_else( usable_em_sent$sentiment == "negative", -1 , 1, 0))

usable_em_sent <- usable_em_sent %>% 
  group_by(Song)%>%
   summarise(net_Sen = sum(sen_value))


p1 <- ggplot(usable_ej_sent, aes( x = reorder(Song, net_Sen), y = net_Sen, fill = net_Sen < 0)) + 
  geom_bar(stat = "identity") + ggtitle("Song sentiment \n by Elton john songs") + theme(plot.title=element_text(face="bold")) +
   xlab("Songs") + ylab("Count")+
    scale_fill_manual(values = c("blue", "red"),
                    labels=c('positive song','negative song'))+
  labs(fill='net sentiment')+
  coord_flip()
p1

p2 <- ggplot(usable_em_sent, aes( x = reorder(Song, net_Sen), y = net_Sen, fill = net_Sen < 0)) + 
  geom_bar(stat = "identity") + ggtitle("Song sentiment \n by Eminem songs") +   theme(plot.title=element_text(face="bold")) +
   xlab("Songs") + ylab("Count")+
    scale_fill_manual(values = c( "red"),
                    labels= c('negative song')) +
  labs(fill='net sentiment')+
  coord_flip()
p2




```

- What differences do you notice between the 2?

ANSWER -> Elton John has songs that have both positive and negative sentiments.
wheras Eminem has songs that are mostly negative.  

- Based on what you might know about some of these songs, do you think the sentiment analysis gives a good indication of the positive of negative tone for each song or not?

ANSWER -> It tallies with the general trend of most songs from Eminem but there are some songs like Lose yourself - which are widely known as motivational song, but due to the wording lose, it might be perceived as negative by the algorithem.

- If you're interested, play around with this data set, see if your favorite artist is in here and see if you can find anything else

Just a reminder: if you have questions, please ask them on Slack.  

#### Links
 
http://www.r-project.org 

http://rmarkdown.rstudio.com/  